Marlowe: financial contracts on blockchain\thanks{This work is part of the Cardano project and is supported by IOHK,~\url{https://iohk.io}}

### 2018-07-26 tex/paper.tex
Give names to participants in escrow contract

Consider the following example of an ``escrow'' contract so that we can explain the motivation 
more concretely. The aim of this contract, written in functional pseudocode in the style of~\cite{PeytonJones:2000} 
involves three participants \haskellinline{1}, \haskellinline{2} and \haskellinline{3}. Participant \haskellinline{1} is 
to pay an amount of money to participant \haskellinline{2} on receipt of goods from her. Participant \haskellinline{1} 
pays the money into escrow controlled by participant \haskellinline{3}.

There are two options for the money: if two participants agree to \haskellinline{pay} it to \haskellinline{2}, that goes 
ahead; if, on the other hand, two of the participants opt to \haskellinline{refund} the money to \haskellinline{1}, that 
is done instead.

@@

\begin{haskellcode}
(When (Or (two_chose 1 2 3 refund)
          (two_chose 1 2 3 pay))
      (Choice (two_chose 1 2 3 pay)
              (Pay 1 2 AvailableMoney)
              redeem_original))
\end{haskellcode}

We discuss this particular example in more detail in Marlowe in Section~\ref{section:example-escrow} below; but it 
already gives us an example of how traditional contracts are fundamentally different from contracts that are meant to 
be run on top of the blockchain. In the traditional model, enforcement of the contract is the responsibility of the 
legal system. If participant \haskellinline{1} does not pay the money into escrow, or participant \haskellinline{3} 
chooses to keep it for herself, then they can be sued for the money (and probably damages), thus providing both legal 
and financial incentives for compliance. On the other hand, in the decentralised blockchain model, where there is no 
central authority, the contract needs to be enforced \emph{by design}.

@@

If the cash is committed successfully and the timeout \haskellinline{end_timeout} is reached, then it is impossible to 
further spend the committed cash, and any unspent funds can be reclaimed by participant \haskellinline{person}.
 This is enforced by the \haskellinline{stepBlock} function, as noted above.

@@

In this section, we show how Marlowe can be used for designing a contract that incentivises saving. We use the 
following example: participant \haskellinline{2} offers 20 ADA to participant \haskellinline{1} in exchange for saving 
100 ADA for 100 blocks. participant \haskellinline{1} will be able to withdraw the savings after block 20, but if that 
option is chosen, then participant \haskellinline{2} will not pay the extra 20 ADA to participant \haskellinline{1}.

We will start from outer to inner primitives. The first thing is to allow participant \haskellinline{1} to commit the 
100 ADA. The primitive says that with identity \haskellinline{IdentCC 1} (this is the identity of the commitment), 
participant \haskellinline{1} can commit \haskellinline{100} ADA before block \haskellinline{10}, and the commitment 
must expire on block \haskellinline{200}. But if the money is not committed before block \haskellinline{10}, then the 
contract will continue as \haskellinline{Null}.

@@

If the commitment is done correctly and on time, we continue as in \haskellinline{step2}. \haskellinline{step2} is 
another commitment, this time by participant \haskellinline{2}. The primitive says that with identity 
\haskellinline{IdentCC 2}, participant \haskellinline{2} can commit \haskellinline{20} ADA before block 
\haskellinline{20}, and the commitment must expire on block \haskellinline{200}. But if the money is not committed 
before block \haskellinline{20}, then the contract will continue as \haskellinline{RedeemCC (IdentCC 1) Null}, which in 
turn gives participant \haskellinline{1} the opportunity to withdraw all the money committed immediately.

@@

If the commitment is done correctly and on time, we continue as in \haskellinline{step3}. \haskellinline{step3}
waits until choice with identity \haskellinline{IdentChoice 1} is made by participant \haskellinline{1} or
until block \haskellinline{100} (whatever happens first). If the choice is made before block \haskellinline{100}, 
it continues as \haskellinline{stepRedeem}, if block \haskellinline{100} comes first, then it continues as 
\haskellinline{step4}.

@@

\haskellinline{step4} pays the \haskellinline{20} ADA committed by participant \haskellinline{2} in the commit with 
identity \haskellinline{IdentPay 1} to participant \haskellinline{1}, if the payment is claimed before block 
\haskellinline{200}. Once the payment is claimed or after block 200, it continues as \haskellinline{stepRedeem}.

@@

The example we use through this section implements an escrow contract, as first introduced in Section~\ref{sec:intro}. 
%
%Let us imagine that participant 
%\haskellinline{1} is buying some item from participant \haskellinline{2}. On one hand, if participant \haskellinline{1} 
%pays before receiving the item, there is a risk that participant \haskellinline{2} will never give \haskellinline{1} the 
%item. On the other hand, if participant \haskellinline{2} gives participant \haskellinline{1} the item before 
%\haskellinline{1} pays,  there is a risk that \haskellinline{1} will never pay.
%
The escrow mechanism allows participant \haskellinline{1} to deposit the money into a contract, in a way that the money 
will only be released when two out of three participants agree on whether participant \haskellinline{2} has indeed given 
participant \haskellinline{1} the item.

The third participant (participant \haskellinline{3}) is supposed to be a 
neutral third party that will decide in case of dispute.
This way, if participants \haskellinline{1} and \haskellinline{2} are honest, they will just agree on the result of the 
transaction and \haskellinline{3} will not need to do anything. If \haskellinline{1} and \haskellinline{2} disagree, 
participant \haskellinline{3} will be able to choose whether the money must go to \haskellinline{1} or to 
\haskellinline{2}.

In our implementation we make things more specific: the money paid for the item is 450 ADA, and it must be committed by 
participant \haskellinline{1} before block \haskellinline{10}; it will be refunded to participant \haskellinline{1} if 
there is no consensus before block \haskellinline{90}.

@@

\begin{haskellcode}
iCC1 :: IdentCC
iCC1 = IdentCC 1
\end{haskellcode}
And an identifier for the payment:
\begin{haskellcode}
iP1 :: IdentPay
iP1 = IdentPay 1
\end{haskellcode}
We can also create a sub-contract that allows the money from the commitment with identifier \haskellinline{IdentCC 1} to 
be redeemed:
\begin{haskellcode}
redeem_original :: Contract
redeem_original = RedeemCC iCC1 Null
\end{haskellcode}
Once redeemed, the contract continues as \haskellinline{Null} since we expect this to be the last thing that is done.
Each participant has a say on who deserves the money: either participant \haskellinline{1} deserves a refund, which we 
represent with the number $0$; or participant \haskellinline{2} deserves a payment, which we represent with the number 
$1$. Because there is only one choice to make per participant, we use the same \haskellinline{IdentChoice} as their 
participant id. We can define a function that returns an observation that is true if and only if
\haskellinline{per} has chosen the number \haskellinline{c} for the choice \haskellinline{IdentChoice per} as follows:
\begin{haskellcode}
chose :: Int -> ConcreteChoice -> Observation
chose per c = PersonChoseThis (IdentChoice per) per c
\end{haskellcode}
Then, we can easily write a function that returns an observation that is \haskellinline{True} if and only if at least 
one of the participants \haskellinline{per} and \haskellinline{per'} has chosen the number \haskellinline{val} as 
follows:
\begin{haskellcode}
one_chose :: Person -> Person -> ConcreteChoice -> Observation
one_chose per per' val = (OrObs (chose per val) (chose per' val)) 
\end{haskellcode}
Building on that, we can now write a function that returns an observation that is \haskellinline{True} if and only if 
at least two out of the three participants \haskellinline{p1}, \haskellinline{p2}, and \haskellinline{p3} have agreed 
in choosing the number \haskellinline{c} as follows:
\begin{haskellcode}
two_chose :: Person -> Person -> Person -> ConcreteChoice -> Observation
two_chose p1 p2 p3 c = OrObs (AndObs (chose p1 c) (one_chose p2 p3 c))
                             (AndObs (chose p2 c) (chose p3 c))
\end{haskellcode}
Finally, we can write the escrow contract, thus:
\begin{haskellcode}
escrow :: Contract
escrow = CommitCash iCC1 1 (ConstMoney 450) 10 100
                    (When (OrObs (two_chose 1 2 3 0)
                                 (two_chose 1 2 3 1))
                          90
                          (Choice (two_chose 1 2 3 1)
                                  (Pay iP1 1 2 (AvailableMoney iCC1) 100
                                       redeem_original)
                                  redeem_original)
                          redeem_original)
                    Null
\end{haskellcode}

The outermost primitive \haskellinline{CommitCash} allows the participant \haskellinline{1} to commit 
\haskellinline{450} ADA before block \haskellinline{10}, with the promise that money will be released on block 
\haskellinline{100} if they are not claimed before that.

@@

If the third option happens, the money is refunded. Otherwise, there is a \haskellinline{Choice} that will 
immediately refund the money to participant \haskellinline{1} unless two out of three chose option \haskellinline{1}; 
in the later case, \haskellinline{Pay} will give participant \haskellinline{2} the opportunity to claim the funds 
available in the commitment with identifier \haskellinline{iCC1} before block \haskellinline{100} (by using the 
identifier \haskellinline{iP1} for the claim). If the funds are not claimed before block \haskellinline{100} they will 
also be refunded to participant \haskellinline{1}.

Reflecting on the example, we can see that using Haskell definitions has made the contract substantially more 
comprehensible. While our current implementation does not do this, it is possible to modify the embedding to support 
more efficient operation, by preserving \emph{sharing} in the host language. For example, if we were to replace the 
repeated  expression 
\haskellinline{two_chose 1 2 3 1} by a \haskellinline{where} clause, 
\begin{haskellcode}
escrow = ...        (When (OrObs (two_chose 1 2 3 0)
                                 chose_refund)
                          90
                          (Choice chose_refund ...
                           ...
         where chose_refund = two_chose 1 2 3 1                 
\end{haskellcode}
then the repeated computation of the expression could be avoided.
%In the next section, we cover the online contract visualisation and interaction tool for Marlowe.

>>>>

Consider the following example of an ``escrow'' contract so that we can explain the motivation 
more concretely. The aim of this contract, written in functional pseudocode in the style of~\cite{PeytonJones:2000} 
involves three participants: \haskellinline{alice}, \haskellinline{bob} and \haskellinline{carol}. \haskellinline{alice} 
is to pay an amount of money to \haskellinline{bob} on receipt of goods from her. \haskellinline{alice} pays the money 
into escrow controlled by \haskellinline{carol}.

There are two options for the money: if two out of the three participants agree to \haskellinline{pay} it to 
\haskellinline{bob}, that goes ahead; if, on the other hand, two of the participants opt to \haskellinline{refund} the 
money to \haskellinline{alice}, that is done instead.

@@

\begin{haskellcode}
(When (Or (two_chose alice bob carol refund)
          (two_chose alice bob carol pay))
      (Choice (two_chose alice bob carol pay)
              (Pay alice bob AvailableMoney)
              redeem_original))
\end{haskellcode}

We discuss this particular example in more detail in Marlowe in Section~\ref{section:example-escrow} below; but it 
already gives us an example of how traditional contracts are fundamentally different from contracts that are meant to 
be run on top of the blockchain. In the traditional model, enforcement of the contract is the responsibility of the 
legal system. If \haskellinline{alice} does not pay the money into escrow, or \haskellinline{carol} chooses to keep it 
for herself, then they can be sued for the money (and probably damages), thus providing both legal and financial 
incentives for compliance. On the other hand, in the decentralised blockchain model, where there is no central 
authority, the contract needs to be enforced \emph{by design}.

@@

If the cash is committed successfully and the timeout \haskellinline{end_timeout} is reached, then it is impossible to 
further spend the committed cash, and any unspent funds can be reclaimed by \haskellinline{person}.
 This is enforced by the \haskellinline{stepBlock} function, as noted above.

@@

In this section, we show how Marlowe can be used for designing a contract that incentivises saving. We use the 
following example: \haskellinline{bob} offers 20 ADA to \haskellinline{alice} in exchange for saving 100 ADA for 100 
blocks. \haskellinline{alice} will be able to withdraw the savings after block 20, but if that option is chosen, then 
\haskellinline{bob} will not pay the extra 20 ADA to participant \haskellinline{alice}.

We will start from outer to inner primitives. The first thing is to allow \haskellinline{alice} to commit the 
100 ADA. The primitive says that with identity \haskellinline{IdentCC 1} (this is the identity of the commitment), 
\haskellinline{alice} can commit \haskellinline{100} ADA before block \haskellinline{10}, and the commitment 
must expire on block \haskellinline{200}. But if the money is not committed before block \haskellinline{10}, then the 
contract will continue as \haskellinline{Null}.

@@

If the commitment is done correctly and on time, we continue as in \haskellinline{step2}. \haskellinline{step2} is 
another commitment, this time by \haskellinline{bob}. The primitive says that with identity 
\haskellinline{IdentCC 2}, \haskellinline{bob} can commit \haskellinline{20} ADA before block 
\haskellinline{20}, and the commitment must expire on block \haskellinline{200}. But if the money is not committed 
before block \haskellinline{20}, then the contract will continue as \haskellinline{RedeemCC (IdentCC 1) Null}, which in 
turn gives \haskellinline{alice} the opportunity to withdraw all the money committed immediately.

@@

If the commitment is done correctly and on time, we continue as in \haskellinline{step3}. \haskellinline{step3}
waits until choice with identity \haskellinline{IdentChoice 1} is made by \haskellinline{alice} or
until block \haskellinline{100} (whatever happens first). If the choice is made before block \haskellinline{100}, 
it continues as \haskellinline{stepRedeem}, if block \haskellinline{100} comes first, then it continues as 
\haskellinline{step4}.

@@

\haskellinline{step4} pays the \haskellinline{20} ADA committed by \haskellinline{bob} in the commit with 
identity \haskellinline{IdentPay 1} to \haskellinline{alice}, if the payment is claimed before block 
\haskellinline{200}. Once the payment is claimed or after block 200, it continues as \haskellinline{stepRedeem}.

@@

The example we use through this section implements an escrow contract, as first introduced in Section~\ref{sec:intro}. 
%
%Let us imagine that participant 
%\haskellinline{alice} is buying some item from \haskellinline{bob}. On one hand, if \haskellinline{alice} 
%pays before receiving the item, there is a risk that \haskellinline{bob} will never give \haskellinline{alice} the 
%item. On the other hand, if \haskellinline{bob} gives \haskellinline{alice} the item before 
%\haskellinline{alice} pays,  there is a risk that \haskellinline{alice} will never pay.
%
The escrow mechanism allows \haskellinline{alice} to deposit the money into a contract, in a way that the 
money will only be released when two out of three participants agree on whether \haskellinline{bob} has indeed 
given \haskellinline{alice} the item.

The escrow participant (\haskellinline{carol}) is supposed to be a neutral third party that will decide in case of 
dispute. This way, if participants \haskellinline{alice} and \haskellinline{bob} are honest, they will just agree on 
the result of the transaction and \haskellinline{carol} will not need to do anything. If \haskellinline{alice} and 
\haskellinline{bob} disagree, \haskellinline{carol} will be able to choose whether the money must go to 
\haskellinline{alice} or to \haskellinline{bob}.

In our implementation we make things more specific: the money paid for the item is 450 ADA, and it must be committed by 
\haskellinline{alice} before block \haskellinline{10}; it will be refunded to participant \haskellinline{alice} if 
there is no consensus before block \haskellinline{90}.

@@

\begin{haskellcode}
iCC1 :: IdentCC
iCC1 = IdentCC 1
\end{haskellcode}
An identifier for the payment:
\begin{haskellcode}
iP1 :: IdentPay
iP1 = IdentPay 1
\end{haskellcode}
And we can create identifiers for all the participants:
\begin{haskellcode}
alice :: Person
alice = 1

bob :: Person
bob = 2

carol :: Person
carol = 3
\end{haskellcode}
We can also create a sub-contract that allows the money from the commitment with identifier \haskellinline{IdentCC 1} 
to be redeemed:
\begin{haskellcode}
redeem_original :: Contract
redeem_original = RedeemCC iCC1 Null
\end{haskellcode}
Once redeemed, the contract continues as \haskellinline{Null} since we expect this to be the last thing that is done.
Each participant has a say on who deserves the money: either \haskellinline{alice} deserves a refund, which 
we represent with the number $0$; or \haskellinline{bob} deserves a payment, which we represent with the 
number $1$. Because there is only one choice to make per participant, we use the same \haskellinline{IdentChoice} as 
their participant id. We can define a function that returns an observation that is true if and only if
\haskellinline{per} has chosen the number \haskellinline{c} for the choice \haskellinline{IdentChoice per} as follows:
\begin{haskellcode}
chose :: Int -> ConcreteChoice -> Observation
chose per c = PersonChoseThis (IdentChoice per) per c
\end{haskellcode}
Then, we can easily write a function that returns an observation that is \haskellinline{True} if and only if at least 
one of the participants \haskellinline{per} and \haskellinline{per'} has chosen the number \haskellinline{val} as 
follows:
\begin{haskellcode}
one_chose :: Person -> Person -> ConcreteChoice -> Observation
one_chose per per' val = (OrObs (chose per val) (chose per' val)) 
\end{haskellcode}
Building on that, we can now write a function that returns an observation that is \haskellinline{True} if and only if 
at least two out of the three participants \haskellinline{p1}, \haskellinline{p2}, and \haskellinline{p3} have agreed 
in choosing the number \haskellinline{c} as follows:
\begin{haskellcode}
two_chose :: Person -> Person -> Person -> ConcreteChoice -> Observation
two_chose p1 p2 p3 c = OrObs (AndObs (chose p1 c) (one_chose p2 p3 c))
                             (AndObs (chose p2 c) (chose p3 c))
\end{haskellcode}
Finally, we can write the escrow contract, thus:
\begin{haskellcode}
escrow :: Contract
escrow = CommitCash iCC1 1 (ConstMoney 450) 10 100
                    (When (OrObs (two_chose alice bob carol 0)
                                 (two_chose alice bob carol 1))
                          90
                          (Choice (two_chose alice bob carol 1)
                                  (Pay iP1 alice bob (AvailableMoney iCC1) 100
                                       redeem_original)
                                  redeem_original)
                          redeem_original)
                    Null
\end{haskellcode}

The outermost primitive \haskellinline{CommitCash} allows the \haskellinline{alice} to commit 
\haskellinline{450} ADA before block \haskellinline{10}, with the promise that money will be released on block 
\haskellinline{100} if they are not claimed before that.

@@

If the third option happens, the money is refunded. Otherwise, there is a \haskellinline{Choice} that will 
immediately refund the money to \haskellinline{alice} unless two out of three chose option 
\haskellinline{1}; 
in the later case, \haskellinline{Pay} will give \haskellinline{bob} the opportunity to claim the funds 
available in the commitment with identifier \haskellinline{iCC1} before block \haskellinline{100} (by using the 
identifier \haskellinline{iP1} for the claim). If the funds are not claimed before block \haskellinline{100} they will 
also be refunded to \haskellinline{alice}.

Reflecting on the example, we can see that using Haskell definitions has made the contract substantially more 
comprehensible. While our current implementation does not do this, it is possible to modify the embedding to support 
more efficient operation, by preserving \emph{sharing} in the host language. For example, if we were to replace the 
repeated  expression 
\haskellinline{two_chose alice bob carol 1} by a \haskellinline{where} clause, 
\begin{haskellcode}
escrow = ...        (When (OrObs (two_chose alice bob carol 0)
                                 chose_refund)
                          90
                          (Choice chose_refund ...
                           ...
         where chose_refund = two_chose alice bob carol 1                 
\end{haskellcode}
then the repeated computation of the expression could be avoided.
%In the next section, we cover the online contract visualisation and interaction tool for Marlowe.
